#Early Stages Using Toy Dataset
###Tuning Not Yet Implemented
###Automation Not Yet Implemented

#Stock Market Data

```{r}
library(ISLR)
names(Smarket)
dim(Smarket)
summary(Smarket)
pairs(Smarket) #Pairwise Scatterplot (Hard to see with several variables)
#cor(Smarket)
cor(Smarket[,-9]) #9th column (Direction) is qualitative, remove it for correlation
attach(Smarket)
plot(Volume)
```

#Logistic Regression

```{r}
glm.fit=glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume,data=Smarket,family=binomial)
summary(glm.fit)
coef(glm.fit)
summary(glm.fit)$coef[,4] #p-values are part 4 of the coefficient summary
glm.probs=predict(glm.fit,type="response")
glm.probs[1:10]
contrasts(Direction)
```

For now, we see 0 means the market will go down and 1 means the market will go up.
Convert predicted probabilities ino class labels.
Create a vector of Up/Down or Yes/No values.
First, Create a vector of "No" or "Down".  All 1250 values will be Down.
Next, when the probability exceeds a certain threshold, we change the value to "Up"
```{r}
glm.pred=rep("Down",1250)
glm.pred[glm.probs>.5]="Up"
```

Create a Confusion Matrix to determine how many observations were correctly labeled.

```{r}
table(glm.pred,Direction)
(507+145)/1250
mean(glm.pred==Direction) #Does the same as above
```

Note: The value above is only the TRAINING Accuracy Rate, since this was performed only on Training Data.

Create a quick train-test split

```{r}
train=(Year<2005)
Smarket.2005=Smarket[!train,]
dim(Smarket.2005)
Direction.2005=Direction[!train]  #Recall, we atttached Smarket

glm.fit=glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume,data=Smarket,family=binomial,subset=train)
glm.probs=predict(glm.fit,Smarket.2005,type="response")
```

Note: We Fit on Train and Predict on Test

```{r}
glm.pred=rep("Down",252) #252 from dim(Smarket.2005)
glm.pred[glm.probs>.5]="Up" #like before, building a vector of Up and Down
table(glm.pred,Direction.2005)

mean(glm.pred==Direction.2005) #percent accurate 
mean(glm.pred!=Direction.2005) #percent error
```


To predict new data, create just predict on a new dataframe
predict(glm.fit,newdata=data.frame(Lag1=c(1.2,1.5),Lag2=c(1.1,-0.8)) ... Volume=c(10,29),type="response")
