#Linear Model Selection
```{r, echo=TRUE}
datafile <- read.csv(file="C:/Users/andrew/Desktop/Boston2.csv",head=TRUE,sep=",")
datafile$X <- NULL
head(datafile)
colnames(datafile)[colnames(datafile)=="medv"] <- "dep_var"
head(datafile)
#dim(datafile)
datafile <- na.omit(datafile)
#dim(datafile)
ncol.less.one <- as.numeric(ncol(datafile) -1)
#(2/3)*nrow(datafile)
ntrain <- floor((2/3)*nrow(datafile))
#nrow(datafile)-floor((2/3)*nrow(datafile))
```


```{r, echo=TRUE}
library(MASS)
library(ISLR)
library(glmnet)
library(boot)
#head(datafile)
#names(datafile)
#lm.fit=lm(dep_var~lstat + age , data=datafile)
#summary(lm.fit)
```

```{r, echo=TRUE}
#cat("Number of Missing Observations: ", max.adjr2, " variables")
#sum(is.na(datafile$dep_var)) #sum of missing values
cat("Dimensions of Data Set:", "datafile")
cat("Rows: ",nrow(datafile)) 
cat("Columns: ", ncol(datafile))
#dim(datafile)
datafile=na.omit(datafile) #Remove any missing values
cat("Dimensions of Data Set:", "datafile", "After Removing Missing Values")
cat("Rows: ",nrow(datafile)) 
cat("Columns: ", ncol(datafile))
```

```{r, echo=TRUE}
#To Remove Missing Values

#dim(Hitters)
#sum(is.na(Hitters))
#Hitters2 = na.omit(Hitters)
#dim(Hitters2)
```

###Regression Model on Full Model
```{r,echo=TRUE}
full.model <- lm(dep_var ~., data = datafile)
summary(full.model)
```
##Coefficient Estimates: 95% Confidence Interval on Full Model
```{r,echo=TRUE}
confint(full.model)
```

```{r, echo=TRUE}
#install.packages("leaps")
library(leaps)
regfit.full=regsubsets(dep_var~.,datafile,nvmax=ncol(datafile)) #nvmax=8 by default
#summary(regfit.full) #shows plot of selected variables for number of variables included
#?regsubsets
reg.summary=summary(regfit.full)
reg.summary
#names(reg.summary) #list of summary statistics available to help select best model
```
```{r, echo=TRUE}
#To see all R-square, etc
#reg.summary$rsq #R-square for example, increases from 1:n
#reg.summary$adjr2
#reg.summary$rss
#reg.summary$cp
#reg.summary$bic

#par(mfrow=c(2,2))
#plot(reg.summary$rss, xlab="Number of Variables", ylab="RSS", type="l") #type l for connects points with lines
#plot(reg.summary$adjr2, xlab="Number of Variables", ylab="Adjusted Rsq", type="l")
```

#How Many Variables for the Best Performance?
```{r, echo=TRUE}
min.rss <- which.min(reg.summary$rss)
cat("Min RSS occurs with ", min.rss, " variables")

max.rsq <- which.max(reg.summary$rsq)
cat("Max Rsq occurs with ", max.rsq, " variables")

max.adjr2 <- which.max(reg.summary$adjr2)
cat("Max Adjusted Rsq occurs with ", max.adjr2, " variables")

min.cp <- which.min(reg.summary$cp)
cat("Min Cp occurs with ", min.cp, " variables")

min.bic <- which.min(reg.summary$bic)
cat("Min BIC occurs with ", min.bic, " variables")

par(mfrow=c(2,2))
plot(reg.summary$rss, xlab="Number of Variables", ylab="RSS", type="l", main=paste("Min RSS:",  min.rss, sep=" "))
points(min.rss,reg.summary$rss[min.rss], col="red", cex=2,pch=20)
#points(12,reg.summary$adjr2[12], col="red", cex=2,pch=20)
#points(which.max(reg.summary$adjr2),reg.summary$adjr2[which.max(reg.summary$adjr2)], col="red", cex=2,pch=20)

#main=paste("Point",  max.adjr2, "to",max.adjr2,"m", sep=" ")
plot(reg.summary$adjr2, xlab="Number of Variables", ylab="Adjusted Rsq", type="l", main=paste("Max AdjRsq:",  max.adjr2, sep=" "))
points(max.adjr2,reg.summary$adjr2[max.adjr2], col="red", cex=2,pch=20)

plot(reg.summary$cp, xlab="Number of Variables", ylab="Cp", type="l", main=paste("Min Cp:",  min.cp, sep=" "))
points(min.cp,reg.summary$cp[min.cp], col="red", cex=2,pch=20)

plot(reg.summary$bic, xlab="Number of Variables", ylab="BIC", type="l", main=paste("Min BIC:",  min.bic, sep=" "))
points(min.bic,reg.summary$bic[min.bic], col="red", cex=2,pch=20)
```

```{r, echo=TRUE, fig.width=6, fig.height=6}
x <- c(1:ncol(datafile))

plot(regfit.full,scale="r2", main="R2 vs Variable Selection")
axis(4, at=x, labels=x, las=2, tck=-0.01)
mtext("Number of Variables",side=4)
abline(a = max.rsq, b = 0, lwd=3, col = "red")

plot(regfit.full,scale="adjr2", main="Adjusted Rsq vs Variable Selection")
axis(4, at=x, labels=x, las=2, tck=-0.01)
mtext("Number of Variables",side=4)
abline(a = max.adjr2, b = 0, lwd=3, col = "red")

plot(regfit.full,scale="Cp", main="Cp vs Variable Selection")
axis(4, at=x, labels=x, las=2, tck=-0.01)
mtext("Number of Variables",side=4)
abline(a = min.cp, b = 0, lwd=3, col = "red")

plot(regfit.full,scale="bic", main="BIC vs Variable Selection")
axis(4, at=x, labels=x, las=2, tck=-0.01)
mtext("Number of Variables",side=4)
abline(a = min.bic, b = 0, lwd=3, col = "red")

#mtext("Top",side=3)
#mtext("Left",side=2)
#mtext("Bottom",side=1)
```

###Coefficients for R-sqr
```{r,echo=TRUE}
coef(regfit.full,max.rsq)
```
###Coefficients for Adj R-sqr
```{r,echo=TRUE}
coef(regfit.full,max.adjr2)
```
###Coefficients for Cp
```{r,echo=TRUE}
coef(regfit.full,min.cp)
```
###Coefficients for BIC
```{r,echo=TRUE}
coef(regfit.full,min.bic)
```


```{r,echo=TRUE}
regfit.fwd=regsubsets(dep_var~.,data=datafile,nvmax=ncol(datafile),method="forward")
summary(regfit.fwd)
```
```{r,echo=TRUE}
regfit.bwd=regsubsets(dep_var~.,data=datafile,nvmax=ncol(datafile),method="backward")
summary(regfit.bwd)
```

#Validation Set Approach

```{r,echo=TRUE}
set.seed(1)
#dim(datafile) #506x15
train=sample(seq(nrow(datafile)),ntrain,replace=FALSE)
#train
regfit.full=regsubsets(dep_var~.,data=datafile[train,],nvmax=ncol.less.one,method="exhaustive")

val.errors=rep(NA,ncol.less.one)
x.test=model.matrix(dep_var~.,data=datafile[-train,])
for(i in 1:ncol.less.one){
  coefi=coef(regfit.full,id=i)
  pred=x.test[,names(coefi)]%*%coefi
  val.errors[i]=mean((datafile$dep_var[-train]-pred)^2)
}

```

```{r, echo=TRUE}
min(val.errors)
max(val.errors)
which.min(val.errors)

plot(sqrt(val.errors),ylab="Root MSE",xlab="Number of Variables",main="Cross Validation",ylim=c(floor(min(sqrt(regfit.full$rss[-1]/ntrain))),ceiling(max(sqrt(regfit.full$rss[-1]/ntrain)))),pch=19,type="b")
points(which.min(val.errors),sqrt(val.errors)[which.min(val.errors)], col="red",pch=19)

#sqrt(regfit.full$rss[-1]/ntrain)
#floor(min(sqrt(regfit.full$rss[-1]/ntrain)))
#ceiling(max(sqrt(regfit.full$rss[-1]/ntrain)))
points(sqrt(regfit.full$rss[-1]/ntrain),col="blue",pch=19,type="b")
legend("topright",legend=c("Training","Validation", "Best"),col=c("blue","black","red"),pch=19)

#a <- names(coef(regfit.full,which.min(val.errors)))
#a
#b <- data.frame(a)
#b  
#------------------------------
coef(regfit.full,which.min(val.errors))

nms <- names(coef(regfit.full,which.min(val.errors))) 
nms <- c(nms,"dep_var")
nms

idx <- match(nms, names(datafile))
#idx <- sort(c(idx-1,idx))
idx <- sort(c(idx))
NewDF <- datafile[,idx]

head(datafile)
head(NewDF)

#setNames(as.list(c(1,12)),names(coef(regfit.full,which.min(val.errors))))
#setNames

#best.model <- lm(dep_var ~ crim + zn + chas + nox + rm + dis + rad + tax + ptratio + black + lstat , data = datafile[train,])
best.model <- lm(dep_var ~., data=NewDF[train,])
#---------------------------------
#datafile[train,]
```

##Linear Regression Model: Parameter Estimates on Training Data for Validation Set Model
```{r,echo=TRUE}
summary(best.model)
```

##Coefficient Estimates: 95% Confidence Interval
```{r,echo=TRUE}
confint(best.model)
```

